### 2016-07-26 paper/paper_GWR.tex
Rendering.

%% bare_jrnl.tex
%% V1.4
%% 2012/12/27
%% by Michael Shell
%% see http://www.michaelshell.org/
%% for current contact information.
%%
%% This is a skeleton file demonstrating the use of IEEEtran.cls
%% (requires IEEEtran.cls version 1.8 or later) with an IEEE journal paper.
%%
%% Support sites:
%% http://www.michaelshell.org/tex/ieeetran/
%% http://www.ctan.org/tex-archive/macros/latex/contrib/IEEEtran/
%% and
%% http://www.ieee.org/

% *** Authors should verify (and, if needed, correct) their LaTeX system  ***
% *** with the testflow diagnostic prior to trusting their LaTeX platform ***
% *** with production work. IEEE's font choices can trigger bugs that do  ***
% *** not appear when using other class files.                            ***
% The testflow support page is at:
% http://www.michaelshell.org/tex/testflow/

%%*************************************************************************
%% Legal Notice:
%% This code is offered as-is without any warranty either expressed or
%% implied; without even the implied warranty of MERCHANTABILITY or
%% FITNESS FOR A PARTICULAR PURPOSE! 
%% User assumes all risk.
%% In no event shall IEEE or any contributor to this code be liable for
%% any damages or losses, including, but not limited to, incidental,
%% consequential, or any other damages, resulting from the use or misuse
%% of any information contained here.
%%
%% All comments are the opinions of their respective authors and are not
%% necessarily endorsed by the IEEE.
%%
%% This work is distributed under the LaTeX Project Public License (LPPL)
%% ( http://www.latex-project.org/ ) version 1.3, and may be freely used,
%% distributed and modified. A copy of the LPPL, version 1.3, is included
%% in the base LaTeX documentation of all distributions of LaTeX released
%% 2003/12/01 or later.
%% Retain all contribution notices and credits.
%% ** Modified files should be clearly indicated as such, including  **
%% ** renaming them and changing author support contact information. **
%%
%% File list of work: IEEEtran.cls, IEEEtran_HOWTO.pdf, bare_adv.tex,
%%                    bare_conf.tex, bare_jrnl.tex, bare_jrnl_compsoc.tex,
%%                    bare_jrnl_transmag.tex
%%*************************************************************************

% Note that the a4paper option is mainly intended so that authors in
% countries using A4 can easily print to A4 and see how their papers will
% look in print - the typesetting of the document will not typically be
% affected with changes in paper size (but the bottom and side margins will).
% Use the testflow package mentioned above to verify correct handling of
% both paper sizes by the user's LaTeX system.
%
% Also note that the "draftcls" or "draftclsnofoot", not "draft", option
% should be used if it is desired that the figures are to be displayed in
% draft mode.
%
\documentclass[journal]{IEEEtran}
%
% If IEEEtran.cls has not been installed into the LaTeX system files,
% manually specify the path to it like:
% \documentclass[journal]{../sty/IEEEtran}

% Some very useful LaTeX packages include:
% (uncomment the ones you want to load)

% *** MISC UTILITY PACKAGES ***
%
%\usepackage{ifpdf}
% Heiko Oberdiek's ifpdf.sty is very useful if you need conditional
% compilation based on whether the output is pdf or dvi.
% usage:
% \ifpdf
%   % pdf code
% \else
%   % dvi code
% \fi
% The latest version of ifpdf.sty can be obtained from:
% http://www.ctan.org/tex-archive/macros/latex/contrib/oberdiek/
% Also, note that IEEEtran.cls V1.7 and later provides a builtin
% \ifCLASSINFOpdf conditional that works the same way.
% When switching from latex to pdflatex and vice-versa, the compiler may
% have to be run twice to clear warning/error messages.

% *** CITATION PACKAGES ***
%
%\usepackage{cite}
% cite.sty was written by Donald Arseneau
% V1.6 and later of IEEEtran pre-defines the format of the cite.sty package
% \cite{} output to follow that of IEEE. Loading the cite package will
% result in citation numbers being automatically sorted and properly
% "compressed/ranged". e.g., [1], [9], [2], [7], [5], [6] without using
% cite.sty will become [1], [2], [5]--[7], [9] using cite.sty. cite.sty's
% \cite will automatically add leading space, if needed. Use cite.sty's
% noadjust option (cite.sty V3.8 and later) if you want to turn this off
% such as if a citation ever needs to be enclosed in parenthesis.
% cite.sty is already installed on most LaTeX systems. Be sure and use
% version 4.0 (2003-05-27) and later if using hyperref.sty. cite.sty does
% not currently provide for hyperlinked citations.
% The latest version can be obtained at:
% http://www.ctan.org/tex-archive/macros/latex/contrib/cite/
% The documentation is contained in the cite.sty file itself.

% *** GRAPHICS RELATED PACKAGES ***
%
\ifCLASSINFOpdf
  \usepackage[pdftex]{graphicx}
  % declare the path(s) where your graphic files are
  \graphicspath{{/home/saf537/Documents/GWR_hedonicP_MDE/Plots}}
  % and their extensions so you won't have to specify these with
  % every instance of \includegraphics
  \DeclareGraphicsExtensions{.pdf,.jpeg,.png,.jpg}
\else
  % or other class option (dvipsone, dvipdf, if not using dvips). graphicx
  % will default to the driver specified in the system graphics.cfg if no
  % driver is specified.
  \usepackage[dvips]{graphicx}
  % declare the path(s) where your graphic files are
  \graphicspath{{/home/saf537/Documents/GWR_hedonicP_MDE/Plots}}
  % and their extensions so you won't have to specify these with
  % every instance of \includegraphics
  \DeclareGraphicsExtensions{.eps,.pdf,.jpeg,.png,.jpg}
\fi
% graphicx was written by David Carlisle and Sebastian Rahtz. It is
% required if you want graphics, photos, etc. graphicx.sty is already
% installed on most LaTeX systems. The latest version and documentation
% can be obtained at: 
% http://www.ctan.org/tex-archive/macros/latex/required/graphics/
% Another good source of documentation is "Using Imported Graphics in
% LaTeX2e" by Keith Reckdahl which can be found at:
% http://www.ctan.org/tex-archive/info/epslatex/
%
% latex, and pdflatex in dvi mode, support graphics in encapsulated
% postscript (.eps) format. pdflatex in pdf mode supports graphics
% in .pdf, .jpeg, .png and .mps (metapost) formats. Users should ensure
% that all non-photo figures use a vector format (.eps, .pdf, .mps) and
% not a bitmapped formats (.jpeg, .png). IEEE frowns on bitmapped formats
% which can result in "jaggedy"/blurry rendering of lines and letters as
% well as large increases in file sizes.
%
% You can find documentation about the pdfTeX application at:
% http://www.tug.org/applications/pdftex

% *** MATH PACKAGES ***
%
%\usepackage[cmex10]{amsmath}
% A popular package from the American Mathematical Society that provides
% many useful and powerful commands for dealing with mathematics. If using
% it, be sure to load this package with the cmex10 option to ensure that
% only type 1 fonts will utilized at all point sizes. Without this option,
% it is possible that some math symbols, particularly those within
% footnotes, will be rendered in bitmap form which will result in a
% document that can not be IEEE Xplore compliant!
%
% Also, note that the amsmath package sets \interdisplaylinepenalty to 10000
% thus preventing page breaks from occurring within multiline equations. Use:
%\interdisplaylinepenalty=2500
% after loading amsmath to restore such page breaks as IEEEtran.cls normally
% does. amsmath.sty is already installed on most LaTeX systems. The latest
% version and documentation can be obtained at:
% http://www.ctan.org/tex-archive/macros/latex/required/amslatex/math/

% *** SPECIALIZED LIST PACKAGES ***
%
%\usepackage{algorithmic}
% algorithmic.sty was written by Peter Williams and Rogerio Brito.
% This package provides an algorithmic environment fo describing algorithms.
% You can use the algorithmic environment in-text or within a figure
% environment to provide for a floating algorithm. Do NOT use the algorithm
% floating environment provided by algorithm.sty (by the same authors) or
% algorithm2e.sty (by Christophe Fiorio) as IEEE does not use dedicated
% algorithm float types and packages that provide these will not provide
% correct IEEE style captions. The latest version and documentation of
% algorithmic.sty can be obtained at:
% http://www.ctan.org/tex-archive/macros/latex/contrib/algorithms/
% There is also a support site at:
% http://algorithms.berlios.de/index.html
% Also of interest may be the (relatively newer and more customizable)
% algorithmicx.sty package by Szasz Janos:
% http://www.ctan.org/tex-archive/macros/latex/contrib/algorithmicx/

% *** ALIGNMENT PACKAGES ***
%
%\usepackage{array}
% Frank Mittelbach's and David Carlisle's array.sty patches and improves
% the standard LaTeX2e array and tabular environments to provide better
% appearance and additional user controls. As the default LaTeX2e table
% generation code is lacking to the point of almost being broken with
% respect to the quality of the end results, all users are strongly
% advised to use an enhanced (at the very least that provided by array.sty)
% set of table tools. array.sty is already installed on most systems. The
% latest version and documentation can be obtained at:
% http://www.ctan.org/tex-archive/macros/latex/required/tools/

% IEEEtran contains the IEEEeqnarray family of commands that can be used to
% generate multiline equations as well as matrices, tables, etc., of high
% quality.

% *** SUBFIGURE PACKAGES ***
%\ifCLASSOPTIONcompsoc
%  \usepackage[caption=false,font=normalsize,labelfont=sf,textfont=sf]{subfig}
%\else
%  \usepackage[caption=false,font=footnotesize]{subfig}
%\fi
% subfig.sty, written by Steven Douglas Cochran, is the modern replacement
% for subfigure.sty, the latter of which is no longer maintained and is
% incompatible with some LaTeX packages including fixltx2e. However,
% subfig.sty requires and automatically loads Axel Sommerfeldt's caption.sty
% which will override IEEEtran.cls' handling of captions and this will result
% in non-IEEE style figure/table captions. To prevent this problem, be sure
% and invoke subfig.sty's "caption=false" package option (available since
% subfig.sty version 1.3, 2005/06/28) as this is will preserve IEEEtran.cls
% handling of captions.
% Note that the Computer Society format requires a larger sans serif font
% than the serif footnote size font used in traditional IEEE formatting
% and thus the need to invoke different subfig.sty package options depending
% on whether compsoc mode has been enabled.
%
% The latest version and documentation of subfig.sty can be obtained at:
% http://www.ctan.org/tex-archive/macros/latex/contrib/subfig/

% *** FLOAT PACKAGES ***
%
%\usepackage{fixltx2e}
% fixltx2e, the successor to the earlier fix2col.sty, was written by
% Frank Mittelbach and David Carlisle. This package corrects a few problems
% in the LaTeX2e kernel, the most notable of which is that in current
% LaTeX2e releases, the ordering of single and double column floats is not
% guaranteed to be preserved. Thus, an unpatched LaTeX2e can allow a
% single column figure to be placed prior to an earlier double column
% figure. The latest version and documentation can be found at:
% http://www.ctan.org/tex-archive/macros/latex/base/

%\usepackage{stfloats}
% stfloats.sty was written by Sigitas Tolusis. This package gives LaTeX2e
% the ability to do double column floats at the bottom of the page as well
% as the top. (e.g., "\begin{figure*}[!b]" is not normally possible in
% LaTeX2e). It also provides a command:
%\fnbelowfloat
% to enable the placement of footnotes below bottom floats (the standard
% LaTeX2e kernel puts them above bottom floats). This is an invasive package
% which rewrites many portions of the LaTeX2e float routines. It may not work
% with other packages that modify the LaTeX2e float routines. The latest
% version and documentation can be obtained at:
% http://www.ctan.org/tex-archive/macros/latex/contrib/sttools/
% Do not use the stfloats baselinefloat ability as IEEE does not allow
% \baselineskip to stretch. Authors submitting work to the IEEE should note
% that IEEE rarely uses double column equations and that authors should try
% to avoid such use. Do not be tempted to use the cuted.sty or midfloat.sty
% packages (also by Sigitas Tolusis) as IEEE does not format its papers in
% such ways.
% Do not attempt to use stfloats with fixltx2e as they are incompatible.
% Instead, use Morten Hogholm'a dblfloatfix which combines the features
% of both fixltx2e and stfloats:
%
% \usepackage{dblfloatfix}
% The latest version can be found at:
% http://www.ctan.org/tex-archive/macros/latex/contrib/dblfloatfix/

%\ifCLASSOPTIONcaptionsoff
%  \usepackage[nomarkers]{endfloat}
% \let\MYoriglatexcaption\caption
% \renewcommand{\caption}[2][\relax]{\MYoriglatexcaption[#2]{#2}}
%\fi
% endfloat.sty was written by James Darrell McCauley, Jeff Goldberg and 
% Axel Sommerfeldt. This package may be useful when used in conjunction with 
% IEEEtran.cls'  captionsoff option. Some IEEE journals/societies require that
% submissions have lists of figures/tables at the end of the paper and that
% figures/tables without any captions are placed on a page by themselves at
% the end of the document. If needed, the draftcls IEEEtran class option or
% \CLASSINPUTbaselinestretch interface can be used to increase the line
% spacing as well. Be sure and use the nomarkers option of endfloat to
% prevent endfloat from "marking" where the figures would have been placed
% in the text. The two hack lines of code above are a slight modification of
% that suggested by in the endfloat docs (section 8.4.1) to ensure that
% the full captions always appear in the list of figures/tables - even if
% the user used the short optional argument of \caption[]{}.
% IEEE papers do not typically make use of \caption[]'s optional argument,
% so this should not be an issue. A similar trick can be used to disable
% captions of packages such as subfig.sty that lack options to turn off
% the subcaptions:
% For subfig.sty:
% \let\MYorigsubfloat\subfloat
% \renewcommand{\subfloat}[2][\relax]{\MYorigsubfloat[]{#2}}
% However, the above trick will not work if both optional arguments of
% the \subfloat command are used. Furthermore, there needs to be a
% description of each subfigure *somewhere* and endfloat does not add
% subfigure captions to its list of figures. Thus, the best approach is to
% avoid the use of subfigure captions (many IEEE journals avoid them anyway)
% and instead reference/explain all the subfigures within the main caption.
% The latest version of endfloat.sty and its documentation can obtained at:
% http://www.ctan.org/tex-archive/macros/latex/contrib/endfloat/
%
% The IEEEtran \ifCLASSOPTIONcaptionsoff conditional can also be used
% later in the document, say, to conditionally put the References on a 
% page by themselves.

% *** PDF, URL AND HYPERLINK PACKAGES ***
%
%\usepackage{url}
% url.sty was written by Donald Arseneau. It provides better support for
% handling and breaking URLs. url.sty is already installed on most LaTeX
% systems. The latest version and documentation can be obtained at:
% http://www.ctan.org/tex-archive/macros/latex/contrib/url/
% Basically, \url{my_url_here}.

% *** Do not adjust lengths that control margins, column widths, etc. ***
% *** Do not use packages that alter fonts (such as pslatex).         ***
% There should be no need to do such things with IEEEtran.cls V1.6 and later.
% (Unless specifically asked to do so by the journal or conference you plan
% to submit to, of course. )

% correct bad hyphenation here
\hyphenation{op-tical net-works semi-conduc-tor}

\begin{document}
%
% paper title
% can use linebreaks \\ within to get better formatting as desired
% Do not put math or special symbols in the title.
\title{Hedonic Price Modelling with Geographically Weighted Regression in Medellín, Colombia}
%
%
% author names and IEEE memberships
% note positions of commas and nonbreaking spaces ( ~ ) LaTeX will not break
% a structure at a ~ so this keeps an author's name from being broken across
% two lines.
% use \thanks{} to gain access to the first footnote area
% a separate \thanks must be used for each paragraph as LaTeX2e's \thanks
% was not built to handle multiple paragraphs
%

\author{Sara~Arango,
        Juan~Carlos~Duque.% <-this % stops a space
\thanks{Sara Arango Center for Urban Science and Progress, \textit{CUSP}. 
New York University, United States.}% <-this % stops a space
\thanks{Group in Spatial Economics, \textit{RiSE} 
Universidad EAFIT, Medellín, Colombia}}

% note the % following the last \IEEEmembership and also \thanks - 
% these prevent an unwanted space from occurring between the last author name
% and the end of the author line. i.e., if you had this:
% 
% \author{....lastname \thanks{...} \thanks{...} }
%                     ^------------^------------^----Do not want these spaces!
%
% a space would be appended to the last name and could cause every name on that
% line to be shifted left slightly. This is one of those "LaTeX things". For
% instance, "\textbf{A} \textbf{B}" will typeset as "A B" not "AB". To get
% "AB" then you have to do: "\textbf{A}\textbf{B}"
% \thanks is no different in this regard, so shield the last } of each \thanks
% that ends a line with a % and do not let a space in before the next \thanks.
% Spaces after \IEEEmembership other than the last one are OK (and needed) as
% you are supposed to have spaces between the names. For what it is worth,
% this is a minor point as most people would not even notice if the said evil
% space somehow managed to creep in.

% The paper headers
%\markboth{Journal of \LaTeX\ Class Files,~Vol.~11, No.~4, December~2012}%
%{Shell \MakeLowercase{\textit{et al.}}: Bare Demo of IEEEtran.cls for Journals}
% The only time the second header will appear is for the odd numbered pages
% after the title page when using the twoside option.
% 
% *** Note that you probably will NOT want to include the author's ***
% *** name in the headers of peer review papers.                   ***
% You can use \ifCLASSOPTIONpeerreview for conditional compilation here if
% you desire.

% If you want to put a publisher's ID mark on the page you can do it like
% this:
%\IEEEpubid{0000--0000/00\$00.00~\copyright~2012 IEEE}
% Remember, if you use this you must call \IEEEpubidadjcol in the second
% column for its text to clear the IEEEpubid mark.

% use for special paper notices
%\IEEEspecialpapernotice{(Invited Paper)}

% make the title area
\maketitle

% As a general rule, do not put math, special symbols or citations
% in the abstract or keywords.
\begin{abstract}
In this paper we use Geographically Weighted Regression to estimate Real Estate Prices with an hedonic pricing approach, accounting for selected contextual and structural variables for city appraisal records of the Medellín, Colombia in 2014, and claiming that this type of analysis can be equally efficient for emergent cities such as Medellín, regardless of the size of their informal markets.
\end{abstract}

% Note that keywords are not normally used for peerreview papers.
\begin{IEEEkeywords}
Hedonic pricing, real estate, Geographically Weighted Regression.
\end{IEEEkeywords}

% For peer review papers, you can put extra information on the cover
% page as needed:
% \ifCLASSOPTIONpeerreview
% \begin{center} \bfseries EDICS Category: 3-BBND \end{center}
% \fi
%
% For peerreview papers, this IEEEtran command inserts a page break and
% creates the second title. It will be ignored for other modes.
\IEEEpeerreviewmaketitle

\section{Introduction}
% The very first letter is a 2 line initial drop letter followed
% by the rest of the first word in caps.
% 
% form to use if the first word consists of a single letter:
% \IEEEPARstart{A}{demo} file is ....
% 
% form to use if you need the single drop letter followed by
% normal text (unknown if ever used by IEEE):
% \IEEEPARstart{A}{}demo file is ....
% 
% Some journals put the first two words in caps:
% \IEEEPARstart{T}{his demo} file is ....
% 
% Here we have the typical use of a "T" for an initial drop letter
% and "HIS" in caps to complete the first word.
\IEEEPARstart{E}{stimating} the prices of real estate is useful and
necessary for taxation, public policy, infrastructure and equality in cities.

\textbf{Hedonic pricing.} Hedonic pricing is an approach that has been naturally linked to real estate as it welcomes the modelling of assets that are usually not valued by traditional pricing strategies, such as natural resources, accesibility, among others.

GWR is a spatial method developed by Fotheringham et al. \cite{fother} to account for the fact that independent variables might have different effects on a dependent variable depending on spatial location. For example, the effect of distance to subway stations in a city in real estate prices may depend on geographical factors that might not be evident from other independent variables.

\textbf{GWR and hedonic pricing.}

Medellín is the second city of Colombia, with more than two and a half million inhabitants and a population of over three million in its metropolitan area. It is a focus of development for the country, and construction plays a major role in its dynamics. \textbf{CITA SOBRE MEDELLIN SIENDO IMPORTANTE EN REAL ESTATE}.

Most of the formal development in Medellín is very uniform in its types. The developers do not have a clear idea of what customers want, and thus they keep building over the same formula that has worked this far. \textit{Universidad EAFIT} funded a project with its Research Group in Spatial Economics (RiSE) to assess the real estate prices in Medellín, in order to seek for what it is that costumers value in housing, and to set foundations for a spin-off with this prpose.

GWR was used as a method for the hedonic pricing estimation of real estate in Medellín using structural variables (those particular to the housing unit itself) and contextual variables (those related to neighbour quality) to predict total price. An important amount of geographical processing had to be done in order to collect, process and standardize this information, typically unavailable for the real estate business to use.

Some questions arose:

\begin{itemize}
\item How successfully can the price per unit area be modeled? Since Area does explain most of the total price, it would be smart to predict square meter prices, as it could be a more robust measure that can overcome how much area shadows other variables in
the estimation of the total price. Predictability would be lost, but capability to understand what the public really want could be gained.

\item What contextual variables can be meaningful to improve the actual estimation of real estate prices in Medellín? 
\item Does the geographical location of a housing unit absorb other important variables in hedonic pricing theory? How does it relate to them? Which can better explain the price?

\end{itemize}

\section{Methodology}

\subsection{Data and variables}

A first important step of this project was to process geographical data from different sources to add to the original data base that was used and analyzed with Universidad EAFIT.

\begin{enumerate}
\item Housing data was obtained from OIME, Medellín’s municipal Real Estate Market Observatory. It can be found online 1 . Initially, it has 2003 observations for new and used houses and apartments across the city of Medellín in the year of 2014 (See Figure \ref{fig_sim}.).

\begin{figure}[!t]
\centering
\includegraphics[width=2.5in]{/home/saf537/Documents/GWR_hedonicP_MDE/Plots/MDE_point}
\caption{Distribution of data points.}
\label{fig_sim}
\end{figure}

These observations come from appraisals for
taxation purposes. They present a reasonable spatial
distribution. Notice that there are not many data
fields across the center of the city (along the middle
line crossing from south to north) because that is a
restricted zone for construction, and the less dense
zone of the city for residential uses.

The analysed dataset comprises the following
variables. The summary data was calculated after
removing atypical values and observations with
missing data (absolute z-scores greater than 3):

\begin{itemize}
\item Stratum. Categorical index that serves as an
indicator for income. Ranges from one to six. In
Spanish, from lower to higher income: 1 – Uno
(90 counts); 2- Dos (253 counts); 3- Tres (479 counts); 4-Cuatro (246 counts); 5- Cinco (179);
6 –Seis (114).

\item Parking. Boolean factor reflecting whether a residential unit has at least one parking space or not. Counts: 1120 False and 241 True.

\item Utility room. Boolean variable indicating individual storage rooms. Counts: 944 False and 417 True.

\item Price. In COP. By November of 2015, One US Dollar costs around three thousand Colombian pesos.

\item Area. In squared meters.

\item Age. In years. Medellín is a relatively new city that has undergone exponential growth especially since the 1960’s, without much
interest in architectural preservation.

\item Spatial coordinates. X and Y coordinates of data points in WGCS 1984 projected system.

\end{itemize}

\item In order to account for contextual variables (those that could indicate added value for relatively similar properties given their distance to natural features, public spaces or transportation networks) data was obtained from:

\begin{itemize}
\item Medellín’s department of Planning Zoning plan 2 (ArcGIS geodatabase). Includes road networks, metro and BRT stations locations, existing parks, among others. A good amount of geographical processing was needed in order to create variables for nearest distance to features for each of the data points, and to extract
the features that were needed in the case of existing green parks and only the main roads in the network.

\item Neighbor density, given by the Life Quality Survey, performed by the city of Medellín. This polygon data was extracted for each of the data points, depending on their geographical location.

May be this is what you are after:

\begin{table}[htb]
\caption{Summary of the newly created variables.}
\label{1234}
\resizebox{\columnwidth}{!}{%
\begin{tabular}{c|c|c|c|c|c}
value & \textbf{Density} & \textbf{Open space} & \textbf{Transit} & \textbf{Downtown }& \textbf{Roads} \\
\hline
mean & 24630 & 0.001 & 0.0113 & 0.040 & 0.002\\
std & 14576 & 0.001 & 0.0077 & 0.019 & 0.002 \\
min & 0 & 0.000 & 0.0006 & 0.001 & 0.000 \\
25\% & 14453 & 0.000 & 0.0055 & 0.027 & 0.001 \\
50\% & 21218 & 0.001 & 0.0099 & 0.039 & 0.001 \\
75\% & 32645 & 0.001 & 0.0153 & 0.052 & 0.003 \\
max & 77837 & 0.04 & 41.2 & 0.085 & 0.013 \\

\end{tabular}
}
\end{table}

\end{itemize}

\end{enumerate}

\subsection{Methods}

\section{Results}

A simple Spearman’s correlation shows an interesting story (See Figure 2). As it will be discussed, most of the total price of the sampled
houses can be predicted by their total area. It can be seen that Area has a relatively higher than usual degree of correlation to other variables compared to the rest of variable to variable relationships,
although the three highest correlations obtained were:

\begin{itemize}
\item Density and Distance to Downtown are relatively highly positively correlated. This is only a demonstration of the known fact that most of Medellín lives in its periphery \ref{cor}.

\begin{figure}[!t]
\centering
\includegraphics[width=2.5in]{/home/saf537/Documents/GWR_hedonicP_MDE/Plots/cor}
\caption{Distribution of data points.}
\label{cor}
\end{figure}

\item Distance to Mass Transit and Distance to Downtown: Mass transit is focused in certain central corridors where most of the people do not actually live. This is an indicator of the lack of accessibility of this system.

\item Distance to roads and distance to mass transit. This is mainly explained by the fact that the mass transport system overlays the main roads network.

\end{itemize}

Figure 3. displays the relationship between price and other critical variables, colored by strata. There seem to be different behaviors for different strata. This will be further explored below.

By comparing the critical variables mentioned below, some regressions were performed, and it was noted that Density creates multicolinearity issues and does not add to the R2 value, although it is found as
meaningful in a simple regression that includes Area. Distance to green areas also was found not to be significant in a simple regression and also creates multicolinearity with unit area. The same behavior was observed for distance to roads and distance to
public transport.

According to this, the extra geographical features that were added, are not useful for increasing prediction analysis in a global model such as ordinary linear regression. As it will be discussed below, the
explanation for this might be that these contextual variables might only be meaningful when local behaviors are taken into account. So far, with just Strata and Area, most of the variance of the price was
already described, and adding more variables only seems to increase the noise and complexity.

The variable “lot area” was also removed for its correlation to the total area.

\textbf{Figure}

\subsubsection*{Regression analysis}

Using a simple feature selection algorithm that ignores the aforementioned discussion about correlation yields the results displayed in \textbf{Table 3}. Only the presence or absence of a Storage room is not significant with a significance level of 95\%. Distance to mass transit presents a p-value higher than the rest,
but still lower than 0.05 for the stated level of confidence. The independent variables have the expected impacts: Age decreases the value, higher strata imply higher priced properties, parking ads to
the value of the unit, distance to mass transit increases the price of the property. This last observation makes sense in a city like Medellín, where the mass transit system is not very efficient and goes through areas that are traditionally congested, loud, polluted, among
others.

The same simple feature selection model for Price per Unit Area yields the results displayed in Table 4. It is worth noticing that a Principal Component Analysis was used for both of the models and did not yield any significant increase in prediction capabilities nor any
evident cluster more evident than what is already indicated by strata.
The relationship between price per squared meter and area is not that evident, this can be shown by the significant yet negative coefficient in the regression in Table 4, and by the plots displayed in Figure 4. This is consistent with the prior belief that modeling the squared meter can be somehow more informative, and, surprisingly, the estimation capability was not lost at large.

The possibilities displayed here need to be more refined in order to create a robust model that can be used for more general cases in Medellín. It is important to notice that, since the city is very segregated itself, Strata can pick many behaviors that otherwise could be attributed to geographical location.

\textbf{TABLES OF REGRESSION RESULTS}

\textbf{Relationship between variables}

\subsubsection*{Geographically weighted regression}

Although the purpose of this exercise was not to develop a robust brand new geographically weighted regression model, the fact that all the contextual variables created for this exercise were not picked up as important for a global regression, an interest to see how a local model would behave under these conditions. ArcGIS’s GWR method
was used only for Area and Green Spaces. Strata is expected to be picked up by the geography itself.

Since inference is complicated in GWR estimation, no assertions can be made, only that the expected added value to green areas according to location has the expected behaviour that is indeed observed
in Figure 7: wealthier areas are more attracted to this feature, whereas poorer areas actually repel it because it might usually mean less access to services.

\textbf{tables}

\textbf{PLOT}

\section{Discussion}

Not necessarily a novel approach, but it is a novel exercise in the context of emerging cities with potentially very obscure sub markets.

It might be worth just modelling subsets of the city.

GWR must be taken and understood with care since it is a very novel technique and validation is harder, still to be more rigorous.

What is the best way to standardize? what is the best way to treat variables?

Feature selection.

\section{Conclusions}

\begin{itemize}
\item A basic clustering approach does broadly reveal that the submarkets that are expected from the understanding of the city are the ones that are revealed by the prices.
\item The effect of the contextual variables used for this analysis is negligible for a global linear model. Their effect might be masked by a Simpson’s paradox like effect, and thus the importance of geographically weighted regression or other local method. This result
might not mean that these variables should be discarded for Geographically Weighted Regression, only that global analyses are not
suitable for increasing prediction of real estate prices to a very high degree.
\item Using the square meter price of housing units can be a useful and more robust measure than total price because of the latter’s sensitivity to total area. This needs to be further explored when a complete GWR model is developed.
\end{itemize}

% An example of a floating figure using the graphicx package.
% Note that \label must occur AFTER (or within) \caption.
% For figures, \caption should occur after the \includegraphics.
% Note that IEEEtran v1.7 and later has special internal code that
% is designed to preserve the operation of \label within \caption
% even when the captionsoff option is in effect. However, because
% of issues like this, it may be the safest practice to put all your
% \label just after \caption rather than within \caption{}.
%
% Reminder: the "draftcls" or "draftclsnofoot", not "draft", class
% option should be used if it is desired that the figures are to be
% displayed while in draft mode.
%
%\begin{figure}[!t]
%\centering
%\includegraphics[width=2.5in]{myfigure}
% where an .eps filename suffix will be assumed under latex, 
% and a .pdf suffix will be assumed for pdflatex; or what has been declared
% via \DeclareGraphicsExtensions.
%\caption{Simulation Results.}
%\label{fig_sim}
%\end{figure}

% Note that IEEE typically puts floats only at the top, even when this
% results in a large percentage of a column being occupied by floats.

% An example of a double column floating figure using two subfigures.
% (The subfig.sty package must be loaded for this to work.)
% The subfigure \label commands are set within each subfloat command,
% and the \label for the overall figure must come after \caption.
% \hfil is used as a separator to get equal spacing.
% Watch out that the combined width of all the subfigures on a 
% line do not exceed the text width or a line break will occur.
%
%\begin{figure*}[!t]
%\centering
%\subfloat[Case I]{\includegraphics[width=2.5in]{box}%
%\label{fig_first_case}}
%\hfil
%\subfloat[Case II]{\includegraphics[width=2.5in]{box}%
%\label{fig_second_case}}
%\caption{Simulation results.}
%\label{fig_sim}
%\end{figure*}
%
% Note that often IEEE papers with subfigures do not employ subfigure
% captions (using the optional argument to \subfloat[]), but instead will
% reference/describe all of them (a), (b), etc., within the main caption.

% An example of a floating table. Note that, for IEEE style tables, the 
% \caption command should come BEFORE the table. Table text will default to
% \footnotesize as IEEE normally uses this smaller font for tables.
% The \label must come after \caption as always.
%
%\begin{table}[!t]
%% increase table row spacing, adjust to taste
%\renewcommand{\arraystretch}{1.3}
% if using array.sty, it might be a good idea to tweak the value of
% \extrarowheight as needed to properly center the text within the cells
%\caption{An Example of a Table}
%\label{table_example}
%\centering
%% Some packages, such as MDW tools, offer better commands for making tables
%% than the plain LaTeX2e tabular which is used here.
%\begin{tabular}{|c||c|}
%\hline
%One & Two\\
%\hline
%Three & Four\\
%\hline
%\end{tabular}
%\end{table}

% Note that IEEE does not put floats in the very first column - or typically
% anywhere on the first page for that matter. Also, in-text middle ("here")
% positioning is not used. Most IEEE journals use top floats exclusively.
% Note that, LaTeX2e, unlike IEEE journals, places footnotes above bottom
% floats. This can be corrected via the \fnbelowfloat command of the
% stfloats package.

\section{Conclusion}
The conclusion goes here.

% if have a single appendix:
%\appendix[Proof of the Zonklar Equations]
% or
%\appendix  % for no appendix heading
% do not use \section anymore after \appendix, only \section*
% is possibly needed

% use appendices with more than one appendix
% then use \section to start each appendix
% you must declare a \section before using any
% \subsection or using \label (\appendices by itself
% starts a section numbered zero.)
%

\appendices
\section{Geographical inputs}

% you can choose not to have a title for an appendix
% if you want by leaving the argument blank
\section{}
Appendix two text goes here.

% use section* for acknowledgement
\section*{Acknowledgment}

The authors would like to thank the City of Medellín's Real Estate Observatory (OIME) in the head of Juan Pablo Barrero.

% Can use something like this to put references on a page
% by themselves when using endfloat and the captionsoff option.
\ifCLASSOPTIONcaptionsoff
  \newpage
\fi

% trigger a \newpage just before the given reference
% number - used to balance the columns on the last page
% adjust value as needed - may need to be readjusted if
% the document is modified later
%\IEEEtriggeratref{8}
% The "triggered" command can be changed if desired:
%\IEEEtriggercmd{\enlargethispage{-5in}}

% references section

% can use a bibliography generated by BibTeX as a .bbl file
% BibTeX documentation can be easily obtained at:
% http://www.ctan.org/tex-archive/biblio/bibtex/contrib/doc/
% The IEEEtran BibTeX style support page is at:
% http://www.michaelshell.org/tex/ieeetran/bibtex/
%\bibliographystyle{IEEEtran}
% argument is your BibTeX string definitions and bibliography database(s)
%\bibliography{IEEEabrv,../bib/paper}
%
% <OR> manually copy in the resultant .bbl file
% set second argument of \begin to the number of references
% (used to reserve space for the reference number labels box)
\begin{thebibliography}{2}

\bibitem{IEEEhowto:kopka}
H.~Kopka and P.~W. Daly, \emph{A Guide to \LaTeX}, 3rd~ed.\hskip 1em plus
  0.5em minus 0.4em\relax Harlow, England: Addison-Wesley, 1999.

\bibitem{fother}
Fotheringham, A. Stewart., Chris Brundson, and Martin Charlton. \emph{Geographically Weighted Regression: The Analysis of Spatially Varying Relationships}. Chichester, West Sussex, England: Wiley, 2002. Print.

\end{thebibliography}

% biography section
% 
% If you have an EPS/PDF photo (graphicx package needed) extra braces are
% needed around the contents of the optional argument to biography to prevent
% the LaTeX parser from getting confused when it sees the complicated
% \includegraphics command within an optional argument. (You could create
% your own custom macro containing the \includegraphics command to make things
% simpler here.)
%\begin{IEEEbiography}[{\includegraphics[width=1in,height=1.25in,clip,keepaspectratio]{mshell}}]{Michael Shell}
% or if you just want to reserve a space for a photo:

% insert where needed to balance the two columns on the last page with
% biographies
%\newpage

% You can push biographies down or up by placing
% a \vfill before or after them. The appropriate
% use of \vfill depends on what kind of text is
% on the last page and whether or not the columns
% are being equalized.

%\vfill

% Can be used to pull up biographies so that the bottom of the last one
% is flush with the other column.
%\enlargethispage{-5in}

% that's all folks
\end{document}